A Survey of Exoskeleton Control (Draft 1.1)

### 2015-09-18 paper.tex
more updates

@@

Can we build an exoskeleton allows the operator to behave naturally?

@@

@@

@@

@@

\subsection{Reference Design: linear hydraulic}

@@

@@

\begin{verbatim}
sensing
 actuator length/angle
 actuator velocity
 actuator force/torque 
 actuator sensing is pre-transmission
  joint backlash
  joint play
  structural flexibility
  actuator and transmission friction (10 Nm in Atlas)
 joint angle
 joint velocity
 joint torque
 IMUs all over
 cameras all over
\end{verbatim}

@@

Can we use energy storage to assist control: charging batteries,
pressurizing accumulators, using springs, ... ?

@@

@@

Terms like impedance and admittance are used, but are often confusing,
as there are two inputs into the exoskeleton (operator-exo forces and
exo-world contact forces), and the output of the
exoskeleton can be exoskeleton actuator forces and torques, or exoskeleton
motion (position, velocity, and acceleration).

@@

@@

@@

@@

@@

@@

A. M. Dollar, H. Herr, Lower extremity exoskeletons and active
orthoses: challenges and state-of-the-art, IEEE
Transactions on Robotics: 24 (2008) 144-158.
(surveys exoskeletons)
http://excedrin.media.mit.edu/wp-content/uploads/sites/3/2013/07/Dollar-2008_Lower-Extremity-Exoskeletons-and-Active-Orthoses-Challenges-and-State-of-the-Art.pdf
\end{verbatim}

@@

\begin{verbatim}
BLEEX: Berkeley Lower Extremity Exoskeleton

@@

Sarcos XOS/WEAR
http://www.army-technology.com/features/featuremilitary-exoskeletons-uncovered-ironman-suits-a-concrete-possibility
http://www.robaid.com/bionics/raytheons-second-generation-exoskeleton-xos-2.htm

@@

@@

MINDWALKER
http://www.3me.tudelft.nl/en/about-the-faculty/departments/biomechanical-engineering/research/dbl-delft-biorobotics-lab/exoskeleton/

@@

>>>>

This could be achieved through active control using feedback of the
operator-exoskeleton forces. Additional gravity, friction, actuator, and
inertial compensation (inverse dynamics) can improve performance.

@@

Can we build an exoskeleton allows the operator to behave naturally?
The operator can feel the exoskeleton, but can overpower it when necessary.

Bypass valves and clutches allow the operator to physically ``take over''
and push the relatively lightweight leg mounted exoskeleton around.

``Passive'' degrees of freedom can have physical or ``virtual'' springs
applied to assist in support (all abduct/adduct DOF, rotational DOF).
These aids can be clutched out or turned off when necessary.
``Virtual'' sprintgs can be modulated.

@@

The exoskeleton can adapt its behavior to the task.
For example, for jumping, the exoskeleton can operate like a pogo 
stick \url{https://www.youtube.com/watch?v=lbp41vWP4o4}, 
and for running it could act like jumping 
stilts \url{https://www.youtube.com/watch?v=9ZOd7yEyhwI},

@@

{\bf EMG}
Electrical activity was measured using surface electrodes,
high-pass filtered at 20 Hz, rectified, low-pass filtered at 6 Hz,
offset by a small value of -0.008, and amplified by a gain of 283
to obtain the desired torque.~\cite{IEEE07139980}

Electromyography (EMG)
signals are extremely noisy and it is not easy to extract a
meaningful data form the raw signal. In addition, it takes a long
time to attach the EMG electrodes to the body and calibrate the
sensors.
\begin{verbatim}
T. Hayashi, H. Kawamoto, and Y. Sankai, "Control method of robot
suit HAL working as operator's muscle using biological and dynamical
information," in Intelligent Robots and Systems, 2005.(IROS 2005).
2005 IEEE/RSJ International Conference on, 2005, pp. 3063-3068.
H. Kawamoto and Y. Sankai, "Power assist system HAL-3 for gait
disorder person," in Computers helping people with special needs, ed:
Springer, 2002, pp. 196-203.
\end{verbatim}

@@

Hip, knee, ankle: At least 200Nm, ideally >400Nm.

@@

\subsection{Reference Design: linear hydraulic with bypass valves}

@@

{\it 140 bar(2,050 psi) and 8 l pm in total. Hydraulic cylinders
are designed to be double-acting at the hip pitch joint and
the knee pitch joint. The hip and knee cylinders are designed
for a maximum thrust of 4 kN and 2 kN, respectively. The
four three-way servo valves(M 200, Star-Hydraulic)}~\cite{IEEE07222598}

@@

\begin{verbatim}
sensing
 actuator length/angle
 actuator velocity
 actuator force/torque 
n-line loadcells(31E-05KNO,
Sensotec) are equipped on each cylinder tube end for the
actuator force.~\cite{IEEE07222598}
 actuator sensing is pre-transmission
  joint backlash
  joint play
  structural flexibility
  actuator and transmission friction (10 Nm in Atlas)
 joint angle
three high precision(17 bit) encoders(DS-25-16, Net-
zer) are used to measure the robots joint angles(hip, knee,
ankle pitch angle).~\cite{IEEE07222598}
 joint velocity
 joint torque
 IMUs all over
 cameras all over
\end{verbatim}

{\it foot sensors, which are comprised
of four one-axis(normal force) loadcells(CBFSB-200, Bong-
shin), are adopted to detect the gait phase. The placement
of each loadcell is chosen from the analysis of GRF for
level walking : Hallux(toe), first Metatarsal(inside), fourth
Metatarsal(outside), and Heel similar to [20], as shown in
Fig. 2(a). The Gait phase detection is simply classified
into two phases by the threshold method using the sum
of the load cells: Single Stance (Left Stance(LS) or Right
Stance(RS)), Double Stance(DS). The result of gait detection
while walking is shown in Fig. 2(b).}~\cite{IEEE07222598}
20: Kong K, Tomizuka M. ”A gait monitoring system based on air pressure
sensors embedded in a shoe,” IEEE/ASME Trans Mechatron, 358 - 370.
2009.

@@

Can we use local energy storage to assist control: charging batteries,
pressurizing accumulators, using springs, ... ?

{\it incorporation of physical
compliance is of importance when it comes to dependability,
low mechanical impedance, locomotion efficiency, inherent
safety and enhanced environmental interaction capabilities}~\cite{IEEE07139975}

@@

stance leg requires a large torque and
relatively low bandwidth during the support period. However,
the swing leg undergoes a large motion, but it only supports
its own weight during the swing period. Therefore, the swing
leg requires a relatively small torque. Thus, a dual-mode
approach is very useful for an exoskeleton robot control.
In this paper, a dual-mode control method using a hydraulic
system is proposed,~\cite{IEEE07222598}

rise from squat, chair.
push large objects out of the way (standing full body push
or seated using legs to push)

@@

One basic distinction is whether the actuators are thought of as position
or force sources. Often this is revealed by the design of the ``low level''
controller. Note that this is often somewhat confusing. Electric motors
are torque sources (low impedance), but when a high gear ratio transmission
is added (such as a harmonic drive) the whole system becomes high impedance
and thus is best thought of as a position (or velocity) source. Therefore,
position control is performed by the low level controller. Hydraulic actuators
are high impedance, but when a force or load sensor is added (or piston differential
oil pressure is used as a force sensor), the whole system becomes low impedance
and is best thought of as a force source. Force (or joint torque) control is
performed by the low level controller. Force or torque can be added to an
electromechanical drive train (electric motor plus gears or motor plus ballscrew)
to make it lower impedance and more like a force source as well.

Terms like impedance and admittance control are used, but are often confusing,
as in controller design one can choose from several possible
inputs into the exoskeleton (exoskeleton positions,
velocities, accelerations, operator-exo contact forces and
exo-world contact forces), and choose from several possible exoskeleton
outputs: exoskeleton actuator forces and torques, exoskeleton
motion (position, velocity, and acceleration), as well as operator-exo
contact forces or exo-world contact forces.

@@

Variants of nonlinear feedback control such as feedback linearization or
sliding mode control are largely ignored. Once the decision to use feedback control
based on a set of observable quantities and with particular outputs is made,
one can try out the various linear and nonlinear feedback control paradigms
to see what works well.

@@

Although impedance control are extensively utilized in rehabilitation
robotics [18], limited number of studies has been focused on
this method for power augmentation [19, 20].
\begin{verbatim}
[18] R. Riener, L. Lunenburger, S. Jezernik, M. Anderschitz, G. Colombo,
and V. Dietz, "Patient-cooperative strategies for robot-aided treadmill
training: first experimental results," Neural Systems and Rehabilitation
Engineering, IEEE Transactions on, vol. 13, pp. 380-394, 2005.
[19] B.-K. Lee, H.-D. Lee, J.-y. Lee, K. Shin, J.-S. Han, and C.-S. Han,
"Development of dynamic model-based controller for upper limb
exoskeleton robot," in Robotics and Automation (ICRA), 2012 IEEE
International Conference on, 2012, pp. 3173-3178.
[20] W. Yu, J. Rosen, and X. Li, "PID admittance control for an upper limb
exoskeleton," in American Control Conference (ACC), 2011, 2011, pp.
1124-1129.
\end{verbatim}

@@

Friction, actuator dynamics, exo-world contact forces.

@@

\subsection{Sensitive Amplicfication Control}

BLEEX:
Sensitivity amplification control (SAC) was firstly
developed by Kazerooni [14]. In this method, the interaction
force from human are regarded as disturbance. The control
system has high sensitivity to disturbances and thus has fast
motion response to the interaction forces. The main drawback
of SAC is unwanted response to any disturbance regardless of
being intentional or non-intentional disturbance. In addition,
the highly accurate dynamic modelling is required in this
method.
\begin{verbatim}
H. Kazerooni, J.-L. Racine, L. Huang, and R. Steger, "On the control of
the berkeley lower extremity exoskeleton (BLEEX)," in Robotics and
Automation, 2005. ICRA 2005. Proceedings of the 2005 IEEE
International Conference on, 2005, pp. 4353-4360.
\end{verbatim}

\subsection{Dual Control Appproach}

{\it The robot utilized the dual-mode control scheme, which is
comprised of the active control for the stance phase and the
passive control (using bypass valves) for the swing phase, to achieve high walking
speed in the swing phase while supporting heavy loads in
the stance phase. To reduce the sudden change of the torque
command at the transition from the swing phase to the stance
phase, a smoothing method is adopted. We also implemented
a pre-transition method to take a foot off quickly for fast
walking by predicting the change from the swing to the stance
in advance.}~\cite{IEEE07222598}

Bypass valves implement what Sarcos calls ``dangle''.

stance: virtual joint torque control method Kazerooni et al.(2005)
When contact location between the wearer and the
exoskeleton is not fixed and difficult to estimate, this method
has been shown to be an effective method to generate the
locomotion for an exoskeleton robot.[3],[4],[21]
H. Kazerooni , Z. Racine , L. Huang and R. Steger ”On the control of
the Berkeley lower extremity exoskeleton (BLEEX),” Proc. IEEE Int.
Conf. Robot. Autom., pp. 4364-4371, 2005.
Racine Jean-Louis Charles ”Control of a Lower Extrmity Exoskeleton
for Human Performance Amplification,” University of California,2003.
582
Xiuxia Yang, Hongchao Zhao, Yi Zhang, Xiaowei Liu ”Carrying
Lower Extreme Exoskeleton Rapid Terminal Sliding-Mode Robust
Control,” Journal of computers, vol.9, No.1,202-208. 2014.

to reduce
sudden changes(command jump) at the phase transitions, a
smoothing method [5],[19] is introduced and a pre-transition
method is used to solve the swing delay due to the internal
pressure(approx 5 bar).
H. Kazerooni , Ryan Steger, Lihua Huang ”Hybrid Control of the
Berkeley Lower Extremity Exoskeleton (BLEEX),” Int. Journal of
Robotics Research, pp. 561-573, 2006.
oonbum Bae, Kyoungchul Kong, Masayoshi Tomizuka ”Gait Phase-
Based Control for a Rotary Series Elastic Actuator Assisting the knee

Transition Control

1) Smoothing Method: During transitions of the gait
phase, discontinuity of the control command torque are
occurred by the different condition for the fixed coordinate
which is on the backpack at the swing phase or the foot at
the stance phase. In this paper, to reduce this sudden change
due to gait phase changes, a smoothing method is proposed
as shown in Fig. 7. An exponential function is considered
as the weighting function for smoothing as shown in (6).
In this case, the weighting is small at the initial stage but
it exponentially converges to one for supporting the load
quickly.

2) Pre-transition Method: The pre-transition method is
that the passive mode is executed in the pre-swing phase
prior to toe off. This dramatically reduces the moving

\subsection{Ground Reaction Force Control}

The ground reaction
forces (GRF) magnitude and direction is used to command the
actuators. In some researches the GRF sensors are used
together with other sensors in the control architectures [17],
while in other researches the control system is based on the
GRF merely [RoboKnee, Honda].
\begin{verbatim}
17: K. Suzuki, G. Mito, H. Kawamoto, Y. Hasegawa, and Y. Sankai,
"Intention-based walking support for paraplegia patients with Robot
Suit HAL," Advanced Robotics, vol. 21, pp. 1441-1469, 2007
\end{verbatim}

@@

H. Lee, W. Kim, J. Han, and C. Han, ”The technical trend of the
exoskeleton robot system for human power assistance, ” International
Journal of Precision Engineering and Manufacturing, vol. 13, pp. 1491-
1497, 2012.

@@

A. M. Dollar, H. Herr, Lower extremity exoskeletons and active
orthoses: challenges and state-of-the-art, IEEE
Transactions on Robotics: 24 (2008) 144-158.
(surveys exoskeletons)
http://excedrin.media.mit.edu/wp-content/uploads/sites/3/2013/07/Dollar-2008_Lower-Extremity-Exoskeletons-and-Active-Orthoses-Challenges-and-State-of-the-Art.pdf

Simone Cinquemani Federico Casolo and Matteo Cocetta. On
active lower limb exoskeletons actuators. ISMA 08, 2008
\end{verbatim}

@@

\begin{verbatim}
This is on the low end of the ballpark: 1.1m/s (4km/hr), 99lb (45kg) payload
IEEE 07222598
2015 IEEE International Conference on Advanced Intelligent Mechatronics (AIM)
July 7-11, 2015. Busan, Korea
Locomotion Control Strategy of Hydraulic Lower Extremity
Exoskeleton Robot
Hongchul Kim, Changhoon Seo, Young June Shin, Jung Kim, and Youn Sik Kang

BLEEX: Berkeley Lower Extremity Exoskeleton

@@

Sarcos XOS/WEAR
http://www.army-technology.com/features/featuremilitary-exoskeletons-uncovered-ironman-suits-a-concrete-possibility
http://www.robaid.com/bionics/raytheons-second-generation-exoskeleton-xos-2.htm
http://www.raytheon.com/newsroom/technology/rtn08_exoskeleton/
http://techcrunch.com/2010/09/27/sarcos-xos2-the-real-life-iron-man/
S. Jacobsen, ”On the development of XOS, a powerful exoskeletal
robot,” IEEEIRSJ IROS, Plenary Talk, 2007.

@@

P.D. Neuhaus, et al., “Design and evaluation of Mina: A robotic
orthosis for paraple ics,” in Rehabilitation Robotics (ICORR), 2011
IEEE International Conference on. 2011. IEEE.

@@

MINDWALKER
key feature is actuation in lateral plane
http://www.3me.tudelft.nl/en/about-the-faculty/departments/biomechanical-engineering/research/dbl-delft-biorobotics-lab/exoskeleton/

@@

”Lightening the Load”, U.S. Naval Research Advisory Committee
(NRAC), 2007.
